---
title: "nocturnal-roost-figures"
author: "Maina Handmaker"
date: "3/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 0. set up
```{r, include = FALSE}
library(lubridate)
library(tidyverse)
library(patchwork)
library(hrbrthemes)
library(ggimage)
library(grid)

deveaux_data <- read_csv("all_count_data.csv")

deveaux_bydate <- deveaux_data %>% 
  mutate(date = make_date(year, month, day))
```

#### 1. Counts in intervals
```{r}
deveaux_30min <- deveaux_bydate %>% 
  dplyr::filter(date %in% as.Date(c("2019-05-04", "2019-05-16", "2019-05-17", "2020-05-04", "2020-05-05"))) %>% 
  group_by(date, interval) %>% 
  summarize(total_count = sum(net.arrival)) %>% 
  # create a new column for cumulative sums
  mutate(cumul = cumsum(total_count)) %>% 
  # just in case because sometimes group_by() does weird things
  ungroup() %>% 
  # create a new column for x-axis in order to use scale_x_datetime
  mutate(time_axis = paste(date, interval) %>% as_datetime())
```

#### 2. Cumulative counts without intervals - so we can plot a "total cumulative count" line calculated using the cumulative counts by each time point, rather than by each half-hour interval used for the bar plot.
```{r}
deveaux_cumul <- deveaux_bydate %>% 
  dplyr::filter(date %in% as.Date(c("2019-05-04", "2019-05-16", "2019-05-17", "2020-05-04", "2020-05-05"))) %>% 
  group_by(date, time) %>% 
  summarize(total_arrival = sum(net.arrival)) %>% 
  mutate(arr_cumul = cumsum(total_arrival)) %>% 
  ungroup() %>% 
  # creating a time_ID: this will come in handy in section 3
  mutate(time_ID = paste(date, time) %>% as_datetime())
```

#### 3. Points on the cumulative count line
Here, we pull the correct count total for the time of civil twilight and high tide on each survey night. These count values are pulled from the `deveaux_cumul` data frame, and will appear later in the code when we use `geom_segment()` and `annotate()` to add vertical lines to our plot to mark the time of civil twilight and high tide. 
```{r}
# function to get the right count (i.e. the right y-axis value)
get_count <- function(datetime) {
  deveaux_cumul %>% 
    filter(time_ID == as_datetime({{ datetime }})) %>% 
    pull(arr_cumul)
}

# civil twilight counts
civ_twi_count_20190504 <- get_count("2019-05-04 20:30:00")
civ_twi_count_20190516 <- get_count("2019-05-16 20:39:00")
civ_twi_count_20190517 <- get_count("2019-05-17 20:40:00")
civ_twi_count_20200504 <- get_count("2020-05-04 20:30:00")
civ_twi_count_20200505 <- get_count("2020-05-05 20:31:00")

# high tide counts
high_tide_count_20190504 <- get_count("2019-05-04 20:39:00")
high_tide_count_20190516 <- get_count("2019-05-16 19:11:00")
high_tide_count_20190517 <- get_count("2019-05-17 20:01:00")
high_tide_count_20200504 <- get_count("2020-05-04 18:25:00")
high_tide_count_20200505 <- get_count("2020-05-05 19:20:00")
```

#### 4. Function  for B&W version of plot for Wader Study publication
```{r}
whim_plot <- function(this_date) {
  ## pre-plotting
  # datetimes for scale_x_datetime
  x_start <- as_datetime(paste(this_date, "16:30:00")) 
  x_end <- as_datetime(paste(this_date, "22:00:00")) 
  
  # times for sunset, sunrise, etc.
  
  civ_twi <- as_datetime(
    paste(
      this_date, (
        if(this_date == "2019-05-04") {
          "20:30:00"
        } else if (this_date == "2019-05-16") {
          "20:39:00"
        } else if (this_date == "2019-05-17") {
          "20:40:00"
        } else if (this_date == "2020-05-04") {
          "20:30:00"
        }  else if (this_date == "2020-05-05") {
            "20:31:00"
        }
      )
    )
  )
  
  high_tide <- as_datetime(
    paste(
      this_date, (
        if(this_date == "2019-05-04") {
          "20:39:00"
        } else if (this_date == "2019-05-16") {
          "19:11:00"
        } else if (this_date == "2019-05-17") {
          "20:01:00"
        } else if (this_date == "2020-05-04") {
          "18:25:00"
        } else if (this_date == "2020-05-05") {
          "19:20:00"
        }
      )
    )
  )
  

  # for points on the cumulative line
  
  hline_yint_civ_twi <- if(this_date == "2019-05-04") {
    civ_twi_count_20190504
  } else if (this_date == "2019-05-16") {
    civ_twi_count_20190516
  } else if (this_date == "2019-05-17") {
    civ_twi_count_20190517
  } else if (this_date == "2020-05-04") {
    civ_twi_count_20200504
  } else if (this_date == "2020-05-05") {
    civ_twi_count_20200505
  }
  
  hline_yint_high_tide <- if(this_date == "2019-05-04") {
    high_tide_count_20190504
  } else if (this_date == "2019-05-16") {
    high_tide_count_20190516
  } else if (this_date == "2019-05-17") {
    high_tide_count_20190517
  } else if (this_date == "2020-05-04") {
    high_tide_count_20200504
  } else if (this_date == "2020-05-05") {
    high_tide_count_20200505
  }
  
  #for plot titles
  plot_title <- if(this_date == "2019-05-04") {
    plot_title <- "4 May 2019"
  } else if (this_date == "2019-05-16") {
    plot_title <- "16 May 2019"
  } else if (this_date == "2019-05-17") {
    plot_title <- "17 May 2019"
  } else if (this_date == "2020-05-04") {
    plot_title <- "4 May 2020"
  } else if (this_date == "2020-05-05") {
    plot_title <- "5 May 2020"
  }
  
  # function to get the right moon phase icon
  
  get_moon <- function(filename) {
    grid::rasterGrob(png::readPNG(filename), interpolate = TRUE)
  }
  
  moon_icon <- get_moon(
    paste(
      if(this_date =="2019-05-04") {
        "4may19_BW.png"
      } else if (this_date == "2019-05-16") {
        "16may19_5may20_BW.png"
      } else if(this_date == "2019-05-17") {
        "17may19_BW.png"
      } else if(this_date == "2020-05-04") {
        "4may20_BW.png"
      } else if(this_date == "2020-05-05") {
        "16may19_5may20_BW.png"
      }
    )
  )

  # new data frame
  new_df <- deveaux_30min %>% 
    filter(date == as.Date({{ this_date }}))
  
  new_cumul_df <- deveaux_cumul %>% 
    filter(date == as.Date({{ this_date }}))
  
  # plot
  fxn_plot <- ggplot(data = new_df) +
    # columns
    geom_col(aes(x = time_axis + minutes(15), y = total_count, 
                 fill = "Whimbrel per half-hour interval"), alpha = 0.8) +
    scale_fill_manual(name = NULL, 
                      values = c("Whimbrel per half-hour interval" = "gray70")) +
    # make the breaks on the x-axis 30 minutes
    scale_x_datetime(breaks = scales::date_breaks("30 mins"), 
                     minor_breaks = scales::date_breaks("30 mins"), 
                     date_labels = "%H:%M",
                     limits = c(x_start, x_end)) +
    # make the breaks on the y-axis 1000
    scale_y_continuous(expand = c(0, 0), limits = c(0, 20000), minor_breaks = seq(from = 0, to = 20000, by = 1000)) +
    
    # cumulative count line
    geom_line(data = new_cumul_df, aes(x = time_ID, y = arr_cumul, 
                                       color = "Total cumulative count", 
                                       linetype = "Total cumulative count",
                                       shape = "Total cumulative count"),
              size = 2.5, show.legend = NA) +
    
    # lines for times of day (vertical)
    geom_segment(data = new_df, aes(x = civ_twi, xend = civ_twi,
                                    y = 0, yend = hline_yint_civ_twi,
                                    color = "Civil twilight", 
                                    linetype = "Civil twilight",
                                    shape = "Civil twilight"),
                 size = 3, show.legend = FALSE, inherit.aes = FALSE) +
    geom_segment(data = new_df, aes(x = high_tide, xend = high_tide,
                                    y = 0, yend = hline_yint_high_tide,
                                    color = "High tide", 
                                    linetype = "High tide",
                                    shape = "High tide"),
                 size = 3, show.legend = FALSE, inherit.aes = FALSE) +
    
    # points on cumulative count line and moon phase icons
    geom_point(aes(x = as_datetime(civ_twi), y = hline_yint_civ_twi,
                   color = "Civil twilight", linetype = "Civil twilight", shape = "Civil twilight"), 
               size = 12, show.legend = NA, inherit.aes = FALSE) +
    geom_point(aes(x = as_datetime(high_tide), y = hline_yint_high_tide,
                   color = "High tide", linetype = "High tide", shape = "High tide"), 
               size = 12, show.legend = NA, inherit.aes = FALSE) +
    annotation_custom(moon_icon, xmin = as_datetime(paste(this_date, "16:30:00")),
                      xmax = as_datetime(paste(this_date, "17:30:00")), ymin = 14000, ymax = 18000) +
    coord_cartesian(clip = "off") +
    
    # colors for lines
    scale_color_manual(name = NULL,
                       values = c("High tide" = "gray38",
                                  "Civil twilight" = "gray20",
                                  "Total cumulative count" = "gray50")) +
    # linetypes for lines
    scale_linetype_manual(name = NULL,
                          values = c("High tide" = "dotted",
                                     "Civil twilight" = "dashed",
                                    "Total cumulative count" = "solid")) +
    # points for points
    scale_shape_manual(name = NULL,
                       values = c("High tide" = 17,
                                  "Civil twilight" = 15,
                                  "Total cumulative count" = NA)) +
    
    
    # appearance
    theme_bw() +
    theme(axis.text.x = element_text(angle = 40, size = 40, face = "bold", hjust = 1.0, color = rep(c("#FFFFFF", "black"), times = 6)),
          axis.text.y = element_text(size = 40, face = "bold", color = "black"),
          axis.title.x = element_text(size = 45, hjust = 1),
          axis.title.y = element_text(size = 45, hjust = 1),
          axis.ticks.x = element_line(size = 1),
          axis.ticks.length.x = unit(.4, "cm"),
          panel.border = element_rect(color = "black", size = 1.5),
          plot.title = element_text(size = 50, hjust = 0.5, face = "bold", margin = margin(b = 15, unit = "pt")),
          plot.margin = unit(c(.5,.5,.5,.5), "cm"),
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          legend.key.width = unit(5, "cm"),
          legend.text = element_text(size = 50)) +
          #legend.position = "bottom",
          #legend.justification = "center") +
    labs(x = "Time",
         y = "Count",
         title = {{ plot_title }})
  
  return(fxn_plot)
}
```



#### 5. Plot each survey night 
```{r, include = FALSE}
# this chunk will throw "Ignoring unknown aesthetics" errors but the plots will still run. This is because linetype, shape and color had to be
# called in the aesthetics for all geoms in order to create the legend we wanted, but R has to ignore the aesthetics that don't exist for some 
# of those geoms - i.e. there is no shape for the cumulative count line
plot_20190504 <- whim_plot("2019-05-04") +
  labs(x = "")
plot_20190516 <- whim_plot("2019-05-16") +
  labs(x = "", y = "") +
  theme(axis.text.y = element_blank())
plot_20190517 <- whim_plot("2019-05-17") +
  labs(y = "") +
  theme(axis.text.y = element_blank())
plot_20200504 <- whim_plot("2020-05-04") +
  labs(x = "")
plot_20200505 <- whim_plot("2020-05-05") +
  labs(y = "") +
  theme(axis.text.y = element_blank())

plot_20190504
plot_20190516
plot_20190517
plot_20200504
plot_20200505
```

#### 6. Patchwork plots together for final figure
```{r, fig.show="hide"}
all_5_nights_BW <- (plot_20190504 + plot_20190516 + plot_20190517 + plot_layout(widths = c(3, 3, 3))) /
  (plot_spacer() + plot_20200504 + plot_20200505 + plot_spacer() + plot_layout(widths = c(1.5, 3.5, 3.5, 1.5))) +
  plot_layout(guides = "collect") & theme(legend.position = "bottom",  legend.box.margin = margin(1, 1, 1, 1, unit = "cm"), legend.box = "vertical")

all_5_nights
```


```{r}
# save 
ggsave("WaderStudy_Fig2_2021-03-13_BW.jpg", all_5_nights_BW, width = 36, height = 20, dpi = 300)
```